---
title: "Danish Kings"
author: "Alexander V. Christiansen"
date: "2025-07-03"
output: html_document
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```


The task here is to load your Danish Monarchs csv into R using the `tidyverse` toolkit, calculate and explore the kings' duration of reign with pipes `%>%` in `dplyr` and plot it over time.

## Load the kings

Make sure to first create an  `.Rproj` workspace with a `data/` folder where you place either your own dataset or the provided `kings.csv` dataset. 

1. Look at the dataset that are you loading and check what its columns are separated by? (hint: open it in plain text editor to see)

List what is the

separator:";" 

2. Create a `kings` object in R with the different functions below and inspect the different outputs.
-  `read.csv()`
-  `read_csv()`
-  `read.csv2()` 
-  `read_csv2()`
```{r eval=TRUE}
# FILL IN THE CODE BELOW and review the outputs

library(tidyverse)

kings1 <- read.csv("data/danish_kings.csv")

kings2 <- read_csv("data/danish_kings.csv")

kings3 <- read.csv2("data/danish_kings.csv")

kings4 <- read_csv2( "data/danish_kings.csv")

class(kings2)
```
Answer: 
1. Which of these functions is a `tidyverse` function? Read data with it below into a `kings` object

 - The read_csv() and read_csv2 functions are tidyverse-functions
 
2. What is the result of running `class()` on the `kings` object created with a tidyverse function.
 - Running this function on the "kings2"-object which is created by using the          following    snippet: "kings2 <- read_csv("data/danish_kings.csv")" we are told that the object is a         dataframe with 54 rows and 18 columns
 
3. How many columns does the object have when created with these different functions?

 - Either 1 or 18 - with 18 being the correct number of columns
 
4. Show the dataset so that we can see how R interprets each column

```{r eval = TRUE}
# COMPLETE THE BLANKS BELOW WITH YOUR CODE, then turn the 'eval' flag in this chunk to TRUE.

#I assign the correct version of my data to the name "kings"
kings <- kings4

#I check the class of the data-frame
class(kings)

#I find the number of columns
ncol(kings)

#I get a "glimpse" of my tibble and the classes of my data
glimpse(kings)

```


## Calculate the duration of reign for all the kings in your table
You can calculate the duration of reign in years with `mutate` function by subtracting the equivalents of your `startReign` from `endReign` columns and writing the result to a new column called `duration`. But first you need to check a few things:

* Is your data messy? Fix it before re-importing to R
* Do your start and end of reign columns contain NAs? Choose the right strategy to deal with them: `na.omit()`, `na.rm=TRUE`, `!is.na()`


Create a new column called  `duration` in the kings dataset, utilizing the `mutate()` function from tidyverse. Check with your group to brainstorm the options.

```{r}
#I assign the new column to the "kings" tibble
kings <- kings %>% 
  mutate(reign_duration = reign_end_year - reign_start_year)
```

## Calculate the average duration of reign for all rulers
Do you remember how to calculate an average on a vector object? If not, review the last two lessons and remember that a column is basically a vector. So you need to subset your `kings` dataset to the `duration` column. If you subset it as a vector you can calculate average on it with `mean()` base-R function. If you subset it as a tibble, you can calculate average on it with `summarize()` tidyverse function. Try both ways!

* You first need to know how to select the relevant `duration` column. What are your options?
* Is your selected `duration` column a tibble or a vector? The `mean()` function can only be run on a vector. The `summarize()` function works on a tibble.
* Are you getting an error that there are characters in your column? Coerce your data to numbers with `as.numeric()`.
* Remember to handle NAs: `mean(X, na.rm=TRUE)`


```{r}
#I find the average using base R
mean(kings$reign_duration, na.rm=TRUE)

#And afterwards the tidyverse functions
kings %>%
  summarize(mean(reign_duration, na.rm = TRUE))

```

## How many and which kings enjoyed a longer-than-average duration of reign?

You have calculated the average duration above. Use it now to `filter()` the `duration` column in  `kings` dataset. Display the result and also  count the resulting rows with `count()`

```{r}
#I define a new tibble only consisting of above-average reigns
kings_abv_avg <- kings %>% 
  filter(reign_duration>=20.59615)

#I count the number of rows in this new tibble
count(kings_abv_avg)

#I use this line again to explicitly showcase my data
kings %>% 
  filter(reign_duration>=20.59615)
```


## How many days did the three longest-ruling monarchs rule? 

* Sort kings by reign `duration` in the descending order. Select the three longest-ruling monarchs with the `slice()` function
* Use `mutate()` to create `Days` column where you calculate the total number of days they ruled
* BONUS: consider the transition year (with 366 days) in your calculation!

```{r}
## Below i am using the slicemax-function to find the three longest ruling monarchs. 

kings %>% 
  slice_max( reign_duration, n = 3)

##Below i have defined new date-columns so that i can use the difftime()-function to calculate the duration of the reigns of the monarchs. For all cases of NA i have used the ifelse()-function, as i just wanted to define those dates as the 1st of january, as the years are known, and it therefore would be stupid to completely neglect the reigns of those kings. 

kings <- kings %>%
  mutate(
    start_date = make_date(reign_start_year, 
                           ifelse(is.na(reign_start_month), 1, reign_start_month), 
                           ifelse(is.na(reign_start_day), 1, reign_start_day)),
    
    end_date = make_date(reign_end_year, 
                         ifelse(is.na(reign_end_month), 1, reign_end_month), 
                         ifelse(is.na(reign_end_day), 1, reign_end_day)),
    
    reign_days = as.numeric(difftime(end_date, start_date, units = "days"))
  )
```


## Challenge: Plot the kings' duration of reign through time

What is the long-term trend in the duration of reign among Danish monarchs? How does it relate to the historical violence trends ? 

* Try to plot the duration of reign column in `ggplot` with `geom_point()` and  `geom_smooth()`
* In order to peg the duration (which is between 1-99) somewhere to the x axis with individual centuries, I recommend creating a new column `midyear` by adding to `startYear` the product of `endYear` minus the `startYear` divided by two (`startYear + (endYear-startYear)/2`).
* Now you can plot the kings dataset, plotting `midyear` along the x axis and `duration` along y axis
* BONUS: add a title, nice axis labels to the plot and make the theme B&W and font bigger to make it nice and legible!

```{r}
#I define a new column of values of the middle year of the reigns
kings <- kings %>% 
  mutate(midyear = (reign_start_year+reign_end_year)/2)

#I plot my function and assign a graph using the local polynomial regression-function geom_smooth. 
kings %>% 
  ggplot(aes(x = midyear , y = reign_duration))+
  geom_point()+
  geom_smooth()+
  theme_bw()+
  labs(
    title = "Duration of Reigns Over Time",
    subtitle = "How long did each monarch reign?",
    x = "Midyear of Reign",
    y = "Reign Duration (years)"
  )
```