---
title: "Letter books - data wrangling"
author: "Alexander V. Christiansen"
date: "2025-05-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The needed libraries for the entire project are activated here, as this Rmd-file is where everything starts to take shape:  
```{r}
library(pdftools)
library(dplyr)
library(readr)
library(qpdf)
library(ggplot2)
library(ggwordcloud)
library(tidytext)
library(tidyverse)
```

Creating a list of the pdf´s, so that these files can be directly called:
```{r}
pdf_files <- list.files(path = "data/brevbøger_opdelt", pattern = "\\.pdf$", full.names = TRUE)
```

Checking to see if R interprets the data correctly: 
```{r}
pdf_info("data/brevbøger_opdelt/1552.pdf")
```

A function is defined, which takes all pdf-files in a defined folder, and turns them into a data frame, with the text in one column and the corresponding year in another column. Afterwards, the function is applied to the letter books:
```{r}
extract_pdf_data <- function(file_path) {

  text <- pdf_text(file_path) %>% paste(collapse = " ") 

  file_name <- basename(file_path)
  year <- gsub("\\.pdf$", "", file_name)
  
  data.frame(Year = year, Content = text, stringsAsFactors = FALSE)
}

pdf_data <- lapply(pdf_files, extract_pdf_data) %>%
  bind_rows()
```
The above code has been partially created by chatgpt, and modified afterwards.

A csv-file is created for easy data management and reproducibility: 
```{r}
write_csv(pdf_data, "data/merged_brevbøger_1551-1660.csv")
```
